---
title: "IMDB Movie Genres"
author: Peter Thompson
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  ioslides_presentation
---
## Introduction

This presentation contains some plots of movie data taken from the [IMDB](http://www.imdb.com/interfaces). Plots are generated using plotly

## libraries
The dplyr and plotly packages are used here

```{r libraries,warning=FALSE,message=FALSE}
library(dplyr)
library(plotly)
```


## Load Genre Data
Will load the genre information. The genres.list file is a text file. A bunch of lines at the beginning should be skipped. The information is also not in a standard table format, so will read the lines individually, parse/clean them using regex (gsub), and then merge the columns into a dataframe.


## Load Genre Data
```{r genres,cache=TRUE,results='hide',warning=FALSE}
lines<-readLines('movies/genres.list')
lines<-lines[384:length(lines)]
Sys.setlocale('LC_ALL','C') 
# remove entries containing curly braces {} - often these are duplicates or movies that were suspended during/before production
badlines<-grepl('\\{.*\\}',lines)
lines<-lines[!badlines]
titles<-gsub('"([^"]+)".*','\\1',lines)
years<-gsub('^.*\\((.+)\\).*','\\1',lines)
years2<-gsub('([0-9]{4,4})/?[A-Z]*','\\1',years)
genres<-gsub('^.*?[ \t]+([A-Za-z-]+)$','\\1',lines)
# get a warning here because the coercion of year to numeric introduces some NAs (this is ok)
genredf<-data.frame(title=titles,year=as.integer(years2),genre=as.factor(genres))
```

## Genres
It might also be interesting to look into ratings information, but the formatting of the ratings.list file makes it difficult to deal with (in R at least).

Instead, will just look at the top 20 genres (i.e., those containing the most movies) over all time

```{r top20,cache=TRUE}
top20<-sort(table(genredf$genre),decreasing=TRUE)[1:20]
#top20
t20list<-names(top20)
slimmed<- genredf %>% filter(genre %in% t20list)
```

## Plotly Plots

Some plots are shown on the next few slides. First a bar chart showing the the number of movies (over all time) belonging to the top 20 genres, and then a scatter plot showing movie numbers per genre vs year.

The code to produce these plots is suppressed here (`echo=FALSE` in the chunk options), because it takes up a lot of space in the presentation slides. The Rmd file can be viewed [here](https://github.com/petethegreat/DataProducts/blob/master/movies.Rmd)


## Top 20 Genres

```{r barchart,cache=TRUE,dependson=top20,warning=FALSE,echo=FALSE}
# bar chart
plot_ly(x=top20,y=names(top20),type='bar',orientation='h',color=~top20)

```

## Genres By Year

```{r plotstuff,dependson='top20',warning=FALSE,cache=TRUE,echo=FALSE}
# plot genres by year
moose2<-slimmed %>% group_by(year,genre) %>% summarize(count=n()) %>% arrange(genre,year)
# head(moose2)
plot_ly(data=moose2,
    x=~year,
    y=~count,
    type='scatter',
    mode='markers+lines',
    color=~genre,
    line=list(width=1)
    ) %>% 
    layout(
        xaxis=list(title='year of movie',range=c(1900,2020)),
        yaxis=list(type='log'))
```

## Notes 
Short movies dominate the movie listings. As these are shorter they probably take much less time and money to produce. Dramas, comedies, and documentaries are also popular.

Not sure why the lines (`mode='markers+lines'`) aren't showing up. Genres tend to follow the same trend. During the 80's and 90's there were more Action movies produced than Romances (because action movies were [awesome](http://www.imdb.com/title/tt0103064/) [then](http://www.imdb.com/title/tt0112864/).



